Eco-evolutionary dynamics, density dependent dispersal, and collective behavior: implications for salmon metapopulation robustness

### 2018-01-20 Manuscript/FinalDraft3/msv10.tex
asym updates

\noindent At low values of density-independent straying the system approaches a fixed point at which both populations persist at equal population size, but as we increase straying, other fixed points are created in which the population sizes are asymmetric (figure \ref{fig:traj}, inset).
The system's underlying symmetry implies that for every asymmetric fixed point there must be another `mirror-image' fixed point in which we find the same population sizes, but where the identities of the populations are reversed.
Asymmetric fixed points appear in bifurcations as a critical value of the straying parameter is crossed. 
As the noise in the system is negligible for the purposes of the bifurcation diagram we can use concepts of deterministic bifurcation theory. 
Based on the Jacobian eigenvalues we conjecture that these bifurcations are fold bifurcations. 
In a generic fold bifurcation of maps, two new fixed points are created, one of which is unstable, while the other is stable. 
In the present case, two of these bifurcations occur at the same time, one which creates fixed points where the first population is dominant in the stable fixed point, whereas the second bifurcation creates the mirror-image fixed points where the second population is dominant in the stable fixed point.

In the asymmetric states the dominant population is well-adapted and reproduces quickly. 
The (small) fraction of this population that strays to the subordinate site constitutes a considerable inflow, such that the population in the subordinate site is not as well adapted, which in turn reduces reproduction and stabilizes the asymmetry. 
In a regime found at low straying rates (regime I) both the symmetric and the asymmetric states are stable fixed points of the system (figure \ref{fig:traj}, dashed lines). 
Which of these fixed points is approached depends on the initial conditions.

As we increase straying, the asymmetric states eventually collide with the stable symmetric state. 
A subcritical pitchfork bifurcation occurs in which the unstable asymmetric states vanishes and the symmetric state is destabilized. 
After this bifurcation the stable asymmetric states are the only attractors. 
We find a wide regime (regime II) where the system will always approach an asymmetric state where one population is suppressed. 
However, if the straying is increased further the imbalance in population sizes becomes harder to maintain. 
Eventually we reach a critical point where the stable asymmetric fixed points become symmetric and collide with the unstable symmetric fixed point. 
The system undergoes a supercritical pitchfork bifurcation, in which the stable asymmetric fixed points vanish while the unstable symmetric fixed point is stabilized. 
After this bifurcation the symmetric fixed point is the only attractor in the system. 
Importantly, we find that increasing the asymmetry in the vital rates of populations between sites does not significantly alter the presence or position of these different regimes (figure \ref{symmetry}).

@@

\begin{figure*}
  \captionsetup{justification=raggedright,
singlelinecheck=false
}
\centering
\includegraphics[width=0.75\textwidth]{fig_rtpe_ddm.pdf}
\caption{
Measures of metapopulation robustness for the density-dependent straying model as a function of individual straying $m_0$ and the strength of collective behaviour $C$ (note the $\rm log_{10}$ scale, including 
a) the Portfolio effect,
b) the time to recovery following near-collapse of both populations,
c) the time to recovery following the extinction of the subordinate population, and
d) the time to recovery following the extinction of the dominant population.
} \label{fig:pert}
\end{figure*}

\begin{figure*}
  \captionsetup{justification=raggedright,
singlelinecheck=false
}
\centering
\includegraphics[width=0.75\textwidth]{fig_hysteresis_ddm.pdf}
\caption{
Alternative stable state regimes I (gray) and II (black) as a function of individual straying $m_0$ and the strength of collective behaviour $C$ (note the $\rm log_{10}$ scale).
Regime I signifies parameter space where there is either 1) an intermediate-density, symmetric, steady state, or 2) an asymmetric dominant/subordinate density.
Regime II signifies parameter space where there is an asymmetric dominant/subordinate steady state density.
The white space to the left (lower values of $m_0$) signifies high-density, symmetric, stable states, and the white space to the right (higher values of $m_0$) signifies low-density, symmetric, steady states.
Relationships are shown for (a) intermediate habitat heterogeneity ($\Delta\theta$), as well as (b) low habitat heterogeneity and (c) high heterogeneity.
} \label{fig:bifurcationsddm}
\end{figure*}

>>>>

\noindent At low values of density-independent straying the system approaches a fixed point at which both populations persist at equal population size, but as we increase straying, other fixed points are created in which the population sizes are asymmetric (figure \ref{fig:traj}, inset).
The system's underlying symmetry implies that for every asymmetric fixed point there must be another `mirror-image' fixed point in which we find the same population sizes, but where the identities of the populations are reversed.
Asymmetric fixed points appear in bifurcations as a critical value of the straying parameter is crossed. 
As the noise in the system is negligible for the purposes of the bifurcation diagram we can use concepts of deterministic bifurcation theory. 
Based on the Jacobian eigenvalues we conjecture that these bifurcations are fold bifurcations. 
In a generic fold bifurcation of maps, two new fixed points are created, one of which is unstable, while the other is stable. 
In the present case, two of these bifurcations occur at the same time, one which creates fixed points where the first population is dominant in the stable fixed point, whereas the second bifurcation creates the mirror-image fixed points where the second population is dominant.

In the asymmetric states the dominant population is well-adapted and reproduces quickly. 
The (small) fraction of this population that strays to the subordinate site constitutes a considerable inflow, such that the population in the subordinate site is not as well adapted, which in turn reduces reproduction and stabilizes the asymmetry. 
In a regime found at low straying rates (regime I) both the symmetric and the asymmetric states are stable fixed points of the system (figure \ref{fig:traj}, purple points). 
Which of these fixed points is approached depends on the initial conditions.

As we increase straying, the asymmetric states eventually collide with the stable symmetric state. 
A subcritical pitchfork bifurcation occurs in which the unstable asymmetric states vanishes and the symmetric state is destabilized. 
After this bifurcation the stable asymmetric states are the only attractors. 
We find a wide regime (regime II) where the system will always approach an asymmetric state where one population is suppressed. 
However, if the straying is increased further the imbalance in population sizes becomes harder to maintain. 
Eventually we reach a critical point where the stable asymmetric fixed points become symmetric and collide with the unstable symmetric fixed point. 
The system undergoes a supercritical pitchfork bifurcation, in which the stable asymmetric fixed points vanish while the unstable symmetric fixed point is stabilized. 
After this bifurcation the symmetric fixed point is the only attractor in the system. 
Importantly, we find that increasing the asymmetry in the vital rates of populations between sites does not significantly alter the presence or position of these different regimes (figure \ref{fig:symmetry}).\\

@@

\begin{figure*}
  \captionsetup{justification=raggedright,
singlelinecheck=false
}
\centering
\includegraphics[width=1\textwidth]{fig_rtpe_ddm.pdf}
\caption{
Measures of metapopulation robustness for the density-dependent straying model as a function of individual straying $m_0$ and the strength of collective behaviour $C$ (note the $\rm log_{10}$ scale, including 
a) the Portfolio effect,
b) the time to recovery following near-collapse of both populations,
c) the time to recovery following the extinction of the subordinate population, and
d) the time to recovery following the extinction of the dominant population.
} \label{fig:pert}
\end{figure*}

\begin{figure*}
  \captionsetup{justification=raggedright,
singlelinecheck=false
}
\centering
\includegraphics[width=1\textwidth]{fig_hysteresis_ddm.pdf}
\caption{
Alternative stable state regimes I (gray) and II (black) as a function of individual straying $m_0$ and the strength of collective behaviour $C$ (note the $\rm log_{10}$ scale).
Regime I signifies parameter space where there is either 1) an intermediate-density, symmetric, steady state, or 2) an asymmetric dominant/subordinate density.
Regime II signifies parameter space where there is an asymmetric dominant/subordinate steady state density.
The white space to the left (lower values of $m_0$) signifies high-density, symmetric, stable states, and the white space to the right (higher values of $m_0$) signifies low-density, symmetric, steady states.
Relationships are shown for (a) intermediate habitat heterogeneity ($\Delta\theta$), as well as (b) low habitat heterogeneity and (c) high heterogeneity.
} \label{fig:bifurcationsddm}
\end{figure*}
